### 2016-03-15 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
unified references to the different relaxations in the numerical results section
also changed align to equation/aligned and made further small changes

\documentclass[journal]{IEEEtran}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{mathtools}
\usepackage{url}
\usepackage{sistyle}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{lemma}[theorem]{Lemma}

\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator{\supp}{supp}

\newcommand{\suchthat}{\,:\,}
\newcommand{\abs}[1]{\lvert{#1}\rvert}
\newcommand{\card}[1]{\lvert{#1}\rvert}
\newcommand{\define}{\coloneqq}
\newcommand{\enifed}{\eqqcolon}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
\newcommand{\Norm}[2]{\lVert{#1}\rVert_{#2}}
\newcommand{\skal}[2]{\langle{#1},{#2}\rangle}
\newcommand{\order}[1]{O\left({#1}\right)}
\newcommand{\T}{^{\top}}
\newcommand{\Tr}{\text{Tr}}
\newcommand{\Rk}{\text{Rk}}
\newcommand{\vect}{\text{vec}}

% sets
\newcommand{\C}{\mathds{C}}
\newcommand{\R}{\mathds{R}}
\newcommand{\Z}{\mathds{Z}}
\newcommand{\ones}{\mathds{1}}

% other commands
\newcommand{\NP}{\text{NP}}
\newcommand{\comment}[1]{\marginpar{\raggedright\tiny {#1}}}

% comma after thousands in numbers
\SIthousandsep{,}

\begin{document}

\title{Computing Exact Restricted Isometry Constants via\\ Mixed Integer Semidefinite Programming}
\author{Tristan~Gally and Marc~E.~Pfetsch
  \thanks{This work was supported by the German Research Foundation (DFG) through Collaborative Research Center 805 and Priority Program 1798.}
  \thanks{The  authors  are  with  the  Research  Group  Optimization,  TU  Darmstadt,
  Darmstadt  64293,  Germany  (e-mail:  gally@mathematik.tu-darmstadt.de;
  pfetsch@mathematik.tu-darmstadt.de).}}

\maketitle

\noindent
\begin{abstract}
  One of the fundamental tasks in compressed sensing is finding the
  sparsest solution to an underdetermined system of linear equations. It is
  well known that although this problem is \NP-hard, under certain
  conditions it can be solved by using a linear program which minimizes the
  1-norm. The restricted isometry property has been one of the key
  conditions in this context. However, computing the best constants for
  this condition is itself \NP-hard. In this paper we propose a
  mixed-integer semidefinite programming approach for computing these
  optimal constants. This also subsumes sparse principal component
  analysis. Computational results with this approach allow to evaluate
  earlier semidefinite relaxations and show that their quality is limited.
\end{abstract}

\section{Introduction}

\IEEEPARstart{F}{inding} a sparsest solution of an underdetermined system of linear
equations is one key problem in compressed sensing and has been
thoroughly investigated in recent years. Given a matrix $A \in \R^{m \times n}$ with $m \leq n$ and 
right-hand side $b \in \R^m$, the core problem is
\begin{align}\label{l0min}
 \min \quad & \Norm{x}{0} \tag{$P_0$}\\
 \text{s.t.} \quad & Ax = b, \nonumber
\end{align}
where $\Norm{x}{0}$ denotes the number of nonzero components of~$x$. This
problem is \NP-hard, see Garey and
Johnson~\cite{GareyJohnson}. One key observation in compressed sensing is
that a solution of~\eqref{l0min} can be found by solving the following
convex optimization problem, if certain conditions are fulfilled:
\begin{align}\label{l1min}
 \min \quad & \Norm{x}{1} \tag{$P_1$}\\
 \text{s.t.} \quad & Ax = b. \nonumber
\end{align}
In fact, this so-called \emph{basis pursuit} problem can be reformulated as
a linear program (LP) and efficient specialized solution algorithms are
available both in theory and practice: as an extremely selected list
consider the articles~\cite{BecBC11,CheDS99,OsbPT00,vdBF08} and the
practical comparison in~\cite{LorPT15}.
% Note: do not write out author names to keep text short at this place.

% long version:
%
% consider the articles by Chen et al.~\cite{CheDS99} (interior point
% algorithm), Osbourne et al.~\cite{OsbPT00} and Malioutov et
% al.~\cite{MalCW05} ($\ell_1$-homotopy), van den Berg et al.~\cite{vdBF08}
% (SPGL1), and Becker et al.~\cite{BecBC11} (NESTA). More information can be
% found in Foucart and Rauhut~\cite{FouR13} and the practical comparison
% in~\cite{LorPT15}.

\section{The Restricted Isometry Property}
\label{sec:RIP}

\noindent
Several conditions in compressed sensing imply that a solution
of~\eqref{l0min} can be \emph{recovered} by solving \eqref{l1min}, in which
case one speaks of $\ell_0$-$\ell_1$-\emph{equivalence}. First results in
this direction were established by Chen, Donoho, and
Saunders~\cite{CheDS99} and Donoho and Huo~\cite{DH01}, with many articles
following. We therefore refer the reader to the excellent book by Foucart
and Rauhut~\cite{FouR13} for further information.

One important condition is the \emph{restricted isometry property} (RIP),
which was first introduced by Cand\`es and Tao~\cite{CT05}.

\begin{definition}\label{ripdef}
  The \emph{restricted isometry property} (RIP) of order~$k$ holds for a
  matrix $A \in \R^{m \times n}$ with constant~$\delta \geq 0$, if
  \begin{equation}\label{rip}
    (1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2
  \end{equation}
  for all $x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}$.
\end{definition}

Since the following results depend on the constant $\delta$ in \eqref{rip} being
small enough, we are generally interested in the smallest constant satisfying
\eqref{rip}.

\begin{definition}\label{ricdef}
  For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
  \eqref{rip},
  \begin{equation}\label{ric}
    \begin{aligned}
      \delta_k \define \argmin_{\delta \geq 0} \big\{&(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\\
      & \text{ for all }x \in \Sigma_k \big\},
    \end{aligned}
  \end{equation}
  is called \emph{restricted isometry constant} (RIC) of order~$k$.
\end{definition}

If the restricted isometry property holds for a sufficiently small
restricted isometry constant, the optimal solutions of~\eqref{l0min}
and~\eqref{l1min} coincide.  The best-known result of this type is probably
the following:

\begin{theorem}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
  Let $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
  exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ and
  $A\tilde{x} = b$, then the optimal solutions of~\eqref{l0min}
  and~\eqref{l1min} coincide.
\end{theorem}

Many results similar to Theorem~\ref{RIPsqrt2} exist: For
example, Foucart~\cite{fou10} relaxed the condition to $\delta_{2k} <
0.4652$ and Foucart and Rauhut~\cite{FouR13} to $\delta_{2k} < 0.6246$.
Moreover, Cai et al.~\cite{CWX10} showed the condition $\delta_k <
0.307$ for Theorem~\ref{RIPsqrt2} to hold.

For this paper, it is useful to take an asymmetric viewpoint and
distinguish between the lower and upper bounds in~\eqref{rip}, as proposed
by Foucart and Lai~\cite{FL09}:

\begin{definition}\label{asymRIPdef}
 The \emph{lower} and \emph{upper restricted isometry constant} are defined as
 \begin{align}
   \alpha_k & \define \argmax_{\alpha \geq 0}\; \big\{ \alpha^2\, \Norm{x}{2}^2
   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\big\}, \label{lowerRIP}\\
   \beta_k & \define \argmin_{\beta \geq 0}\; \big\{ \Norm{Ax}{2}^2 \leq \beta^2\,
   \Norm{x}{2}^2,\; \text{ for all }x \in \Sigma_k\big\},\label{upperRIP}
 \end{align}
 respectively. The \emph{restricted isometry ratio} (RIR) is
 defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$ for $\alpha_k \neq 0$.
\end{definition}

Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
in the following sense: if the matrix~$A$ satisfies the RIP of order $k$
with RIC $\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
\delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
$\beta_k$, respectively, then this implies a restricted isometry constant
of $\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ in~\eqref{rip}.
So $A$ has a small RIC if and only if it has a large lower and a small upper 
RIC. But since in~\eqref{rip} usually only one of the two inequalities will 
be sharp, the asymmetric version yields more information about~$A$.

For the asymmetric version of the RIP in Definition~\ref{asymRIPdef}, one
can again show $\ell_0$-$\ell_1$-equivalence:

\begin{theorem}[Foucart and Lai~\cite{FL09}]\label{asymRIPtheorem}
  Let $A$ satisfy $\gamma_{2k} \leq 4 \sqrt{2} - 3 \approx 2.6569$. If
  there exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ such
  that $A\tilde{x} = b$, then the optimal solutions of~\eqref{l0min}
  and~\eqref{l1min} coincide.
\end{theorem}

One advantage
of Definition~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
are invariant under scaling of $Ax = b$. The condition of
Theorem~\ref{RIPsqrt2} holding for~$A x = b$ does not imply the same 
for the scaled equation
$\lambda Ax = \lambda b$, $\lambda \in \R \setminus \{0\}$. In Definition~\ref{asymRIPdef}, however, we can
use $\tilde{\alpha}_k = \lambda \alpha_k$, $\tilde{\beta}_k = \lambda
\beta_k$, and therefore $\tilde{\gamma}_k = \lambda^2 \alpha_k^2 /
(\lambda^2 \beta_k^2) = \gamma_k$.

Furthermore, we note that the recovery results given in this section can also be derived 
for the \emph{denoising} case, where instead of~\eqref{l0min} one considers
\begin{equation}\label{l0minDenoising}\tag{$P_0^\delta$}
 \begin{aligned}
  \min \quad & \Norm{x}{0} \\
  \text{s.t.} \quad & \Norm{Ax - b}{2} \leq \delta, 
 \end{aligned}
\end{equation}
for some $\delta \geq 0$; the $\ell_1$-problem~\eqref{l1min} is changed
similarly. For more details we refer the reader to~\cite{FouR13}.

\subsection{Contribution of this Work}

\noindent
There are several types of matrices~$A$ for which it is possible
to show analytically that the RIP holds, for instance random
matrices with high probability, see, e.g., Baraniuk et al.~\cite{BDDW08}.
However, it turns out that computing the restricted isometry
constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. 
Note the unfortunate implication that it is hard to check for a
concrete random matrix whether it satisfies the RIP, since the theoretical
results hold asymptotically and with respect to high probability.

This motivated d'Aspremont et al.~\cite{Asp08,Asp07} to propose several
relaxations based on semidefinite programming (SDP) to compute upper bounds on the restricted isometry
constants; small upper bounds suffice to guarantee
$\ell_0$-$\ell_1$-equivalence between~\eqref{l0min} and~\eqref{l1min}, but
there exist cases in which the bounds are too weak to give a guarantee.

The relaxations in \cite{Asp08,Asp07} were originally derived for the
connected \emph{sparse principal component analysis} (SPCA). Principal
component analysis (PCA) finds orthogonal directions maximizing the 
variance given in matrix $A$, solving problems
\[
\max \{ \Norm{A x}{2}^2 \suchthat \Norm{x}{2} \leq 1\}
\]
along the way. For SPCA, one further adds the condition that $\Norm{x}{0}
\leq k$ with the goal to reduce the dependency on small components. Note
that this is then equivalent to computing the upper RIC and it is \NP-hard
in the strong sense~\cite{PT14}. SPCA was
originally proposed by Zuo et al.~\cite{SPCA}. Other authors like Moghaddam et al.
\cite{sbSPCA} also proposed computing a sparse maximal eigenvalue for SPCA.

The goal of this paper is to compute optimal restricted isometry constants
via mixed-integer semidefinite programming (MISDP). To this end, we will
introduce an MISDP with a rank-1 constraint. In order to utilize this
MISDP, we will show that the rank constraint can be relaxed without
changing the optimal value. We can then apply a MISDP solver.

In our computational results we show that the resulting problems are quite
hard to solve, even for matrices of smaller size. However, this is true for
the SDP relaxations of d'Aspremont et al.\ as well, and it allows for an
evaluation of the quality of these bounds.  In fact, these bounds turn out
to be quite tight or weak, depending on the instance. This complements the
theoretical approximation guarantees developed by d'Aspremont et
al. in~\cite{AspBG14}.

In the following, we start by reviewing the SDP-relaxations of~\cite{Asp08}
and~\cite{Asp07} in Section~\ref{sec:SDPrelax}.  Afterwards, in
Section~\ref{sec:MISDP}, we construct a mixed integer semidefinite
program and show that it produces exact restricted isometry
constants. Finally, in Section~\ref{sec:numerical_results}, we present
numerical results and comparisons with the SDP relaxations in \cite{Asp08}
and \cite{Asp07}.

\section{SDP Relaxations}
\label{sec:SDPrelax}

\noindent
In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed integer semidefinite program we want to 
propose and will also be checked against in the section on numerical
results.

The idea is that the optimal constant $\alpha_k^2$ in \eqref{lowerRIP} can be computed
by the non-convex quadratic optimization problem
\begin{equation}\label{QP}\tag{QP}
 \begin{aligned}
  \min \quad & \Norm{Ax}{2}^2  \\
  \text{s.t.} \quad & \Norm{x}{2}^2 = 1,  \\
  & \Norm{x}{0} \leq k. 
 \end{aligned}
\end{equation}
The corresponding maximization problem allows to compute $\beta_k^2$. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, first introduced by Goemans and Williamson~\cite{GW95} 
for the max-cut problem. The idea is to use new matrix variables $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where, as usual, 
$X \succeq 0$ means that $X$ is symmetric (shortly written $X \in S_n$) and positive semidefinite. The objective and the normalization-constraint can then be expressed as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
\Tr(A\T A X)$ and $1 = \Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$.

In the next step, the inequality $\Norm{x}{0} \leq k$ is then substituted by the weaker constraint $\ones\T
\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}}\, \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank-1 case, where $\vect(X)$ is a vector
in $\R^{n^2}$ consisting of all entries of $X$. 
Relaxing the non-convex rank constraint finally leads to the following SDP:
\begin{equation}\label{Asp07}\tag{A1}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & \ones\T \abs{X} \ones \leq k, \\
  & X \succeq 0. \nonumber
 \end{aligned}
\end{equation}

Another SDP relaxation was proposed by d'Aspremont et al. in \cite{Asp08}, where
\begin{equation}\label{phi}
  \phi(\rho) = \max \big\{ x\T \Sigma\, x - \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1 \big\},
\end{equation}
originally stemming from sparse principal component analysis, is approximately solved for a symmetric positive semidefinite matrix $\Sigma$. For
$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrangian Relaxation as
 \begin{align*}
  \beta_k^2 & = \max \big\{ x\T A\T A x \suchthat \Norm{x}{2} \leq 1,\; \Norm{x}{0} \leq k \big\} \\
	    & \leq \inf_{\rho \geq 0} \max \big\{ x\T A\T A x - \rho 
            (\Norm{x}{0} - k) \suchthat \Norm{x}{2} \leq 1 \big\} \\
	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho\, k. 
 \end{align*}

As one way to solve \eqref{phi} the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
\begin{equation}\label{Asp08Primal}\tag{A2-Primal}
 \begin{aligned}
  \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n,
 \end{aligned}
\end{equation}
for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
square-root of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
\begin{equation}\label{Asp08Dual}\tag{A2-Dual}
 \begin{aligned}
  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \\
  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n \\
  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n,
 \end{aligned}
\end{equation}
which has only half as many variables as \eqref{Asp08Primal} and one fewer linear and semidefinite constraint. Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
optimal solution for \eqref{phi} will always be $x=0$.

As shown in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
\begin{equation}\tag{A2}\label{Asp08}
 \alpha_k^2 \geq \sup_{\rho \geq 0}\; \psi(\rho) - \rho\, k, 
\end{equation}
with 
\begin{equation*}
  \psi(\rho) = \min \big\{ x\T \Sigma\, x + \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1\big\}.
\end{equation*}
In this case problem \eqref{Asp08Primal} or equivalently \eqref{Asp08Dual} has to
be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
to make $\Sigma$ positive semidefinite again. If $\Psi$ is the
corresponding optimal objective, this implies a lower bound on $\alpha_k$
in\comment{Ist $M\,I$ statt $M$ korrekt? \\ $\Rightarrow$ nein, sollten alles Skalare sein, $\Psi$ demenstrechend optimaler Zielfunktionswert statt L\"osung (habe ich jetzt oben angepasst)}
\begin{equation*}
  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \Psi) - \rho\, k.
\end{equation*}
In \cite{AspBG14} explicit bounds on the gap between \eqref{Asp08Primal} and the penalized problem \eqref{phi} were derived using randomization arguments.

%To compute the parameter $\rho$, the authors of \cite{Asp08} propose to minimize over the interval
% \begin{equation*}
%  \max_{i \notin I} (b_i\T x)^2 \leq \rho \leq  \min_{i \in I} (b_i\T x)^2
% \end{equation*}
% the convex function
% \begin{equation*}
%  \text{gap}(\rho) = \lambda_{\max} \left( \sum_{i=1}^n Y_i \right) -
%  \sum_{i \in I} ((b_i\T x)^2 - \rho),
% \end{equation*}
% where $I$ is a given sparsity pattern, found by some heuristic for example, $x$ the largest eigenvector of $\sum_{i \in I} b_i b_i\T$ and
% \begin{equation*}
%  Y_i = \begin{cases}\frac{B_ixx\T B_i}{x\T B_ix}, & \text{for } i \in I \\
%    \max\{0, \rho \frac{b_i\T b_i - \rho}{\rho - (b_i\T x)^2} \} \frac{(I-xx\T)b_ib_i\T(I-xx\T)}{\Norm{(I-xx\T)b_i}{2}^2}, & \text{for } i \in I^c\end{cases}. 
%  %x in the second case sill as above?
% \end{equation*}

\section{An MISDP for the RIP}
\label{sec:MISDP}

\noindent
In this section we propose a MISDP formulation to compute the optimal
constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
\eqref{upperRIP}, respectively. We will again start with the non-convex quadratic problem~\eqref{QP}.
For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$, we
again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
The cardinality-constraint in \eqref{QP} can equivalently be written using
binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i
\leq k$. We can thus equivalently rewrite \eqref{QP} as
\begin{equation}\label{Rk1MISDP}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
  & \sum_{i=1}^n z_i \leq k,\\
  & \Rk(X) = 1, \\
  & X \succeq 0, \\
  & z \in \{0,1\}^n.
 \end{aligned}
\end{equation}
We then relax the non-convex rank constraint to arrive at the following
MISDP formulation:
\begin{equation}\label{MISDP}\tag{MISDP}
 \begin{aligned}
  \min \quad & \Tr(A\T A X) \\
  \text{s.t.} \quad & \Tr(X) = 1, \\
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
  & \sum_{i=1}^n z_i \leq k, \\
  & X \succeq 0, \\
  & z \in \{0,1\}^n.
 \end{aligned}
\end{equation}
In the following, we will show that \eqref{MISDP} actually produces the
exact value of $\alpha_k$. We will use the following result about the
relationship between the rank of solutions of semidefinite programs and the
dimension of corresponding faces of the feasible set~$P$, i.e., subsets $F
\subseteq P$ such that $x \in F$, $y, z \in P$, and $x = \tfrac{1}{2} (y + z)$ imply
$y,z \in F$:

\begin{theorem}[Pataki~\cite{pat98}]\label{patakiLemma}
  Let $X \in F$, where $F$ is a face of
  \begin{equation*}
    P \define \{X \in S_n \suchthat X \succeq 0,\; \Tr(A_i X) = b_i\; \text{ for } i = 1, \dots, \tilde{m}\}
  \end{equation*}
  for symmetric matrices $A_i \in S_n$, $i = 1, \dots, \tilde{m}$. Then
  \begin{equation*}
    \tfrac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq \tilde{m} + dim(F).
  \end{equation*}
\end{theorem}

Using this result on the projection of~\eqref{MISDP} onto the $X$ variables, yields the following.

\begin{theorem}\label{Rk1thm}
  For every $A \in \R^{m \times n}$ and $k \in \Z_{>0}$, there exists an optimal solution $(X^\star, z^\star)$ of \eqref{MISDP} with $\Rk(X^\star) = 1$. 
\end{theorem}

\begin{proof}
  First note that because of the constraint $\Tr(X) = 1$,
  Problem~\eqref{MISDP} is bounded. Moreover, $X_{11} = 1$, $X_{ij} = 0$
  for all $i,j \neq 1$, $z_1 = 1$, $z_i = 0$ for all $i = 2, \dots, n$ is
  feasible for $k > 0$. Thus, an optimal solution to~\eqref{MISDP} exists.

  Let $(X^\star, z^\star)$ be an optimal solution to \eqref{MISDP} with
  $\ell \define \sum_{i=1}^n z^\star_i \leq k$. Let $T \define \{i_1, i_2,
  \dots, i_\ell\}$ be the support of $z^\star$, and define $\check{A} =
  (A_{i_1}, A_{i_2}, \dots, A_{i_\ell})$ to be the submatrix of~$A$ formed by
  the columns indexed by~$T$. Then consider
  \begin{equation}\label{Proj}
   \begin{aligned}
    \min \quad & \Tr(\check{A}\T \check{A} \check{X}), \\
    \text{s.t.} \quad & \Tr(\check{X}) = 1, \\
    & \check{X} \succeq 0,\; \check{X} \in S_\ell.
   \end{aligned}
  \end{equation}

  Let $\check{X}$ be the $\ell \times \ell$ submatrix of $X^\star$ with all
  rows and columns outside of $T$ removed. Then $\check{X}$ is feasible for
  \eqref{Proj} with the same objective value, since we only removed zero
  rows and columns of $X^\star$ and $(A\T A) X^\star$, which do not
  influence the trace or the positive semidefiniteness.

  On the other hand, we can lift any solution $\check{X}$ of~\eqref{Proj}
  to a point $\hat{X} \in S_n$ by extending with zeros. Defining
  $\hat{z}_i$ to be 1 if and only if $i \in T$, yields a feasible point
  $(\hat{X}, \hat{z})$ for \eqref{MISDP} with identical objective value,
  since $\Tr(\check{X}) = 1$ and $\check{X} \in S_\ell$ imply $0 \leq
  \check{X}_{jj} = \hat{X}_{i_j i_j} \leq 1$ for $j = 1, \dots,
  \ell$. Therefore, $\hat{X}_{ij} \leq \hat{z}_j$ by diagonal dominance of
  semidefinite matrices.

  Thus, the optimal objective values of \eqref{Proj} and \eqref{MISDP}
  agree and $\Rk(\hat{X}) = \Rk(\check{X})$.  Therefore, it
  suffices to show that \eqref{Proj} has an optimal solution of rank one.

  Let $\check{X}$ be an extreme point of \eqref{Proj} (since the set of
  optimal points of~\eqref{Proj} is nonempty, convex, and compact, such a
  point exists, see~\cite[Corollary 18.5.1]{Roc70}). Defining the face $F =
  \{\check{X}\}$ and applying Theorem~\ref{patakiLemma} on \eqref{Proj}
  gives
  \begin{equation*}
    \tfrac{1}{2} \Rk(\check{X}) \cdot (\Rk(\check{X})+1) \leq \tilde{m} + dim(F) = 1 + 0 = 1,
  \end{equation*}
  since there are $\tilde{m} = 1$ linear constraints. Since $\check{X} = 0$
  is infeasible for $\Tr(X) = 1$, it follows that $\Rk(\check{X}) = 1$,
  which can then be lifted to an optimal solution of \eqref{MISDP} of rank
  one.
\end{proof}

Theorem~\ref{Rk1thm} guarantees that \eqref{MISDP} always has a solution of
rank one. Therefore, the optimal values of \eqref{MISDP} and \eqref{QP}
agree. We can therefore compute the optimal constant $\alpha_k$ for the RIP
by solving \eqref{MISDP}. Since the same argumentation also holds for the
maximization problem for $\beta_k$, this allows us to compute the
restricted isometry constant and the restricted isometry ratio by
mixed-integer semidefinite programming.

Theorem~\ref{Rk1thm} does not guarantee that any
found solution will have rank one, only that the optimal objective
values agree. This suffices for our application of
computing the RIP constants $\alpha_k$ and $\beta_k$. For other
applications like SPCA, an extreme solution can be generated with an
algorithm proposed by Pataki~\cite{coneLP}.

In order to successfully apply branch and bound to \eqref{MISDP}, it is
important to use strong semidefinite relaxations for fractional $z$
variables. To this end, we can use the following strengthening of the
constraints $-z_j \leq X_{ij} \leq z_j$.

\begin{lemma}\label{HalfConstraint} 
  The inequalities
  \begin{equation}\label{eq:HalfConstraint}
    -\tfrac{1}{2} z_j \leq X_{ij} \leq \tfrac{1}{2} z_j,
  \end{equation}
  for $i, j \in \{1, \dots, n\}$, are valid for~\eqref{MISDP}.
\end{lemma}

\begin{proof}
  Let $X \succeq 0$ satisfy $\Tr(X)=1$. Then $X_{ii}+X_{jj}\leq 1$ and
  nonnegativity of the diagonal entries imply $X_{ii}X_{jj}\leq
  \tfrac{1}{4}$.  Since all minors of $X$ need to be nonnegative, it
  follows that
  \begin{equation*}
    X_{ij}^2 \leq  X_{ii} X_{jj} \leq \tfrac{1}{4}.
  \end{equation*}
  Therefore $- \tfrac{1}{2} \leq X_{ij} \leq \tfrac{1}{2}$, which shows
  validity of~\eqref{eq:HalfConstraint}.
\end{proof}

\section{Numerical Results}
\label{sec:numerical_results}

\noindent
In this section we demonstrate the applicability of the proposed MISDP-formulation and use it to
assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
For solving the mixed-integer semidefinite programs we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
branch and bound framework SCIP 3.2.1 \cite{SCIP} with 
interior-point SDP-solvers, in our case SDPA~7.3.8~\cite{SDPA6,SDPA7}.

The continuous SDPs are solved using the command-line version of
SDPA~7.3.8. We compute the left- and right-hand restricted isometry
constant $\alpha_k$ and $\beta_k$ either exactly using~\eqref{MISDP} or
produce corresponding bounds via~\eqref{Asp07} and~\eqref{Asp08}. In
the latter case, we solve the SDP~\eqref{Asp08Dual} 15 times for $\rho$ between~$0$ and the maximum
diagonal entry of $A\T A$. This is the same approach as in \cite{Asp08},
except that we reduced the number of problems from 25 to 15, as this did
not seem to influence the quality of the solutions much in our preliminary
experiments but brought the solving times more in line with the other
approaches.

We use a testset consisting of 63 matrices of seven 
different types including band matrices, 
with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen 
uniformly in $\{0,1\}$, Gaussian-distributed matrices and rank one 
matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
we also use three more with theoretical results on their restricted isometry
constants. These include matrices with 
$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finally with distribution
\begin{equation}\nonumber
  A_{ij} = \begin{cases}
    + \sqrt{3/m} & \text{with probability } \frac{1}{6}, \\
    0 & \text{with probability } \frac{2}{3}, \\
    - \sqrt{3/m} & \text{with probability } \frac{1}{6}.
  \end{cases}
\end{equation}
For each type we used \mbox{MATLAB 8.3.0} to randomly generate nine
matrices of three different sizes with rows ranging from 15 to 30, columns
from 25 to 40 and order $k$ between three and five.

For the latter three kinds of matrices, it was shown by Baraniuk et al.~\cite{BDDW08} that for $n \rightarrow \infty$, 
sufficiently small $k$, and given $\delta$, they satisfy the restricted isometry property of order $k$  
with probability converging exponentially to one. The needed proportion between $n$ and $k$ is so small, however,
that we cannot expect the generated matrices in small dimensions to satisfy the RIP.

The tests were performed on a Linux cluster with Intel i3 CPUs with 3.2GHz, 4MB cache, and 8GB memory running
Linux. Each computation was performed single-threaded with a single process running on each computer and with 
a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For the 
Problem~\eqref{Asp08}, most of the larger instance were aborted because of the memory limit of 8GB. These are
counted with the maximum allowed time of four hours.

In Table~\ref{lhsRhsGap} we compare the relative gap between the bounds on $\alpha_k$ and $\beta_k$ computed by the two heuristics
and the exact values given by~\eqref{MISDP}. We present arithmetic means over all of the nine matrices for the given type that
could be solved within the time- and memory-limit. Note that we omitted the rank one matrices for the left-hand side since these satisfy
$\alpha_k = 0$ in all cases, so that the trivial lower bound would already be optimal.

For the left-hand-side of the RIP, $\alpha_k$, both relaxations did not perform
too well. Especially~\eqref{Asp08} regularly returned worse solutions than the
trivial lower bound of zero. Relaxation~\eqref{Asp07} performed slightly better, but the bounds
were still not nearly good enough to have any chance of proving $\ell_0$-$\ell_1$-equivalence. The relatively
bad performance of both relaxations for the left-hand side might not be too surprising though, since both relaxations
were originally designed for sparse principal component analysis, which corresponds to the right-hand side
of the RIP.

\input{tables/lhsRhsGap}

For the right-hand side, i.e., $\beta_k$,  both relaxations
perform much better. Problem~\eqref{Asp07} constantly produces bounds within 10 to 30\,\% of the optimal
solution, with only slightly worse results for binary and rank one matrices. The formulation~\eqref{Asp08}
shows bigger fluctuations, but produces better results for five of the seven matrix types, while performing even
worse for binary and rank one matrices. For band matrices, however, the average gap was below 2\,\%, the best 
result for any matrix type among both relaxations.

For reporting some the solving times, we will use the \emph{shifted
  geometric mean} to decrease the influence of easy instances, see
Achterberg~\cite{SCIP} for more details. The shifted geometric
mean of values $x_1$, \dots, $x_n$ is computed as
\begin{equation*}
  \Big( \prod_{i=1}^n (x_i + s)\Big)^{1/n} - s,
\end{equation*}
where we used a shift of $s=10$.

The solving times in seconds for the left- and right-hand side of the RIP are given in
Table~\ref{lhsRhsTime}. The relaxation~\eqref{Asp07}
is very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
cannot compete with these solving times, although for band-matrices, the difference is relatively
small. The Lagrangian Relaxation~\eqref{Asp08}, however, takes more time to produce bounds than is needed
by~\eqref{MISDP} to compute exact solutions. The big difference is also caused by the instances running into the 
memory limit, but even without those, the algorithm still takes slightly longer than~\eqref{MISDP}.

\input{tables/lhsRhsTime}

\section{Conclusion}

\noindent
This paper shows that exact values of restricted isometry constants can be
computed via mixed-integer semidefinite programming. The computed bounds
significantly improve the bounds computed by the semidefinite
relaxations~\eqref{Asp07} and~\eqref{Asp08}
proposed in~\cite{Asp07} and \cite{Asp08}, respectively. In conclusion, the relaxation~
\eqref{Asp07} is fast to compute, but produces a gap of arround 90\,\% for
$\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, 
\eqref{Asp08} provides better bounds for $\beta_k$ for most types of 
matrices, but its running time is too large
to be used in practice. Moreover, directly solving~\eqref{MISDP} is often
faster and produces the exact value. Consequently, 
\eqref{Asp08} should not be used in practice.

Still, the matrices currently handable by~\eqref{MISDP} are small. For the
future, the hope is that with the advancement of solving techniques for
mixed-integer semidefinite programs, the sizes of the matrices can be
significantly increased. The components to be improved include cutting
planes, primal heuristics, and branching rules.

\bibliographystyle{abbrv}
\bibliography{Paperbib}
\end{document}