### 2022-06-10 tex/ConflictAnalysis.tex
1st version of conflict analysis description

\documentclass[10pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[american]{babel}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[lined,ruled,noend,linesnumbered]{algorithm2e}
\usepackage[color=red!50!blue!50,textsize=footnotesize,textwidth=4cm]{todonotes}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=blue]{hyperref}
% \usepackage[tmargin=2.5cm,bmargin=2.5cm,lmargin=2.8cm,rmargin=2.8cm]{geometry}
\usepackage{cleveref}
\usepackage{xspace}

% operators
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\aff}{aff}
\DeclareMathOperator{\lin}{lin}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\relint}{relint}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\Diag}{Diag}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\tr}{tr}
\renewcommand{\Re}{\text{Re}}

% macros
\newcommand{\suchthat}{\,:\,}
\newcommand{\abs}[1]{\lvert{#1}\rvert}
\newcommand{\card}[1]{\lvert{#1}\rvert}
\newcommand{\ceil}[1]{\lceil{#1}\rceil}
\newcommand{\floor}[1]{\lfloor{#1}\rfloor}
\newcommand{\define}{\coloneqq}
\newcommand{\enifed}{\eqqcolon}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
\newcommand{\Norm}[2]{\lVert{#1}\rVert_{#2}}
\newcommand{\skal}[2]{\langle{#1},{#2}\rangle}
\newcommand{\order}[1]{O\left({#1}\right)}
\newcommand{\T}{^{\top}}

% sets
\newcommand{\R}{\mathds{R}}
\newcommand{\Z}{\mathds{Z}}
\newcommand{\Q}{\mathds{Q}}
\newcommand{\N}{\mathds{N}}
\newcommand{\C}{\mathds{C}}
\newcommand{\E}{\mathds{E}}
\newcommand{\ones}{\mathds{1}}
\renewcommand{\P}{\mathds{P}}
\renewcommand{\E}{\mathds{E}}

% environments
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{example}[theorem]{Example}
\newtheorem{question}[theorem]{Question}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}

% adjust itemize
\setlist[itemize]{topsep=0.5ex,partopsep=0ex,parsep=0ex,itemsep=0.5ex}

% other defines
\newcommand{\MP}[1]{\todo{#1}}
\newcommand{\MPin}[1]{\todo[inline]{#1}}

% title
\title{Conflict Analysis for Mixed-integer SDPs}
\author{Marc E. Pfetsch\thanks{Department of Mathematics, TU Darmstadt, Germany}}
\date{June 2022}

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\begin{document}

\maketitle

% -------------------------------------------------------------------------
\section{Introduction}

We consider mixed-integer SDPs of the following form:
\begin{equation}\label{MISDP}
  \begin{aligned}
    \inf \quad & b\T y \\
    \text{s.t.} \quad & \sum_{j=1}^m A^j\, y_j - A^0 \succeq 0, \\
    & D y \geq d,\\
    & \ell \leq y \leq u,\\
    & y_i \in \Z,\quad i \in I.
  \end{aligned}
\end{equation}
In each node, we solve the following relaxations with local bounds $\ell$
and $u$:
\begin{equation}\label{SDP-D}
  \begin{aligned}
    \inf \quad & b\T y \\
    \text{s.t.} \quad & \sum_{j=1}^m A^j\, y_j - A^0 \succeq 0, \\
    & D y \geq d,\\
    & \ell \leq y \leq u,
  \end{aligned}
\end{equation}
where for simplicity of exposition, we assume that all entries of $\ell$
and $u$ are finite.

The general idea of conflict analysis is to generate globally valid
constraints as follows. Let any $\hat{X} \succeq 0$ and $\hat{z} \geq 0$ be
given. Then because $A(y) \define \sum_{j=1}^m A^j\, y_j - A^0$ is positive
semidefinite for all feasible $y \in \R^m$, we have
$\skal{A(y)}{\hat{X}} \geq 0$.  Similarly, the aggregated inequalities
$\hat{z}\T D \geq \hat{z}\T d$ hold. Together this yields:
\begin{align}
  & \skal{A(y)}{\hat{X}} + \hat{z}\T D \geq \hat{z}\T d\notag\\
  \Leftrightarrow\quad
  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D + \geq
    \skal{A^0}{\hat{X}} + \hat{z}\T d,\label{eq:ConflictCut}
\end{align}
which is valid for all feasible $y$. Clearly,
Inequality~\eqref{eq:ConflictCut} is redundant for continuous $y$, but it
might allow to cut off certain mixed-integer solutions~$y$. In the implementation,
we only propagate bounds using~\eqref{eq:ConflictCut}.

More generally, we can also use the variable bounds $\ell \leq y$ and
$y \leq u$. If $\hat{r}^\ell$, $\hat{r}^u \geq 0$ are the corresponding
multiliers, this yields:
\begin{align}
  & \skal{A(y)}{\hat{X}} + \hat{z}\T D + (\hat{r}^\ell)\T y - (\hat{r}^u)\T y
    \geq \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u\notag\\
  \Leftrightarrow
  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D +
    (\hat{r}^\ell)\T y - (\hat{r}^u)\T y \geq \skal{A^0}{\hat{X}} +
    \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u,\label{eq:ConflictCut2}
\end{align}
which is again valid for all feasible $y$.

% -------------------------------------------------------------------------
\section{Generation of Conflict Constraints}

Let $X$ be the dual variables for the SDP-constraint $A(y) \succeq 0$, $z$
be the dual variables for $Dy \geq d$, and $r^\ell$ and $r^u$ be the dual
variables for the bound constraints $y \geq \ell$ and $-y \geq -u$,
respectively. Then the corresponding primal problem for~\eqref{SDP-D} is:
\begin{equation}\label{SDP-P}
  \begin{aligned}
     \sup \quad & \skal{A^0}{X} + z\T d + \ell\T r^\ell - u\T r^u\\
     \text{s.t.} \quad & \skal{A^j}{X} + D\T z + r^\ell_j - r^u_j = b_j && \forall \, j \in [m],\\
     & X \succeq 0,\; z, \; r^\ell,\; r^u \geq 0.
  \end{aligned}
\end{equation}
Any feasible solution can be used to construct~\eqref{eq:ConflictCut}
or~\eqref{eq:ConflictCut2}.

Similarly, a primal ray $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$,
i.e.,
\begin{equation}\label{eq:ray}
\begin{aligned}
  & \skal{A^j}{\hat{X}} + D\T \hat{z} + \hat{r}^\ell_j - \hat{r}^u_j = 0 && \forall\; j \in [m],\\
  & \skal{A^0}{\hat{X}} + d\T \hat{z} + \ell\T \hat{r}^\ell - u\T \hat{r}^u > 0,
\end{aligned}
\end{equation}
yields an inequality as in~\eqref{eq:ConflictCut}
or~\eqref{eq:ConflictCut2}.

If $\ell \leq u$ and~\eqref{eq:ray} has a feasible solution
$(\tilde{X}, \tilde{r}^\ell, \tilde{r}^u)$, then
$(\hat{X}, \hat{r}^\ell, \hat{r}^u)$ with $\hat{X} = \tilde{X}$ and
$\hat{r}^\ell_j = - \min\,\{\skal{A^j}{\tilde{X}}, 0\} \geq 0$ and
$\hat{r}^u_j = \max\,\{\skal{A^j}{\tilde{X}}, 0\} \geq 0$ is feasible
for~\eqref{eq:ray}. Indeed, it is clear that the equations
of~\eqref{eq:ray} are satisfied. Therefore consider the contribution to
$\ell\T \hat{r}^\ell - u\T \hat{r}^u$ in component $j$:
\begin{align*}
  \ell_j \hat{r}^\ell_j - u_j \hat{r}^u_j & = - \ell_j\, \min\,\{\skal{A^j}{\tilde{X}}, 0\} - u_j\, \max\,\{\skal{A^j}{\tilde{X}}, 0\}\\
  & = - \ell_j\, \min\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\} - u_j\, \max\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\}\\
  & = \begin{cases}
    - \ell_j\, (\tilde{r}^u_j - \tilde{r}^\ell_j) & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
    - u_j\, (\tilde{r}^u_j - \tilde{r}^\ell_j) & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
  \end{cases}\\
  & = \begin{cases}
    - \ell_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
    - u_j\, \tilde{r}^u_j + u_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
  \end{cases}\\
  & \geq \begin{cases}
    - u_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if }  \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
    - u_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
  \end{cases}\\
  & = \ell_j \tilde{r}^\ell_j - u_j \tilde{r}^u_j.
\end{align*}
Therefore the inequality in~\eqref{eq:ray} holds as well.

This implies that
\begin{align*}
  0 & < \skal{A^0}{\hat{X}} + \ell\T \hat{r}^\ell - u\T \hat{r}^u\\
  & = \skal{A^0}{\hat{X}} + \sum_{j=1}^m \ell_j \hat{r}^\ell_j -
    \sum_{j=1}^m u_j \hat{r}^u_j\\
  & = \skal{A^0}{\hat{X}} - \sum_{j: \skal{A^j}{\hat{X}} < 0} \ell_j \skal{A^j}{\hat{X}} -
    \sum_{j: \skal{A^j}{\hat{X}} > 0} u_j \skal{A^j}{\hat{X}}.
\end{align*}
This implies
\begin{align*}
  \sum_{j: \skal{A^j}{\hat{X}} < 0} \ell_j \skal{A^j}{\hat{X}} +
    \sum_{j: \skal{A^j}{\hat{X}} > 0} u_j \skal{A^j}{\hat{X}} < \skal{A^0}{\hat{X}}.
\end{align*}
However, any feasible solution $\hat{y}$ satisfies $A(\hat{y}) \succeq 0$ and $\ell
\leq \hat{y} \leq u$ and therefore
\begin{align*}
  0 & \leq \skal{A(\hat{y})}{\hat{X}} =
      \sum_{j=1}^m \skal{A^j}{\hat{X}}\, \hat{y}_j + \sum_{j=1}^m
      \skal{A^j}{\hat{X}}\, \hat{y}_j - \skal{A^0}{\hat{X}}\\
    & \leq \sum_{j: \skal{A^j}{\hat{X}} < 0} \skal{A^j}{\hat{X}}\, \ell_j +
    \sum_{j: \skal{A^j}{\hat{X}} > 0} \skal{A^j}{\hat{X}} u_j -
      \skal{A^0}{\hat{X}},
\end{align*}
a contradition. This implies that the valid linear inequality
$\skal{A(\hat{y})}{\hat{X}} \geq 0$ shows infeasibility with respect to the
local bounds.

% -------------------------------------------------------------------------
\section{Infeasibility Proof Analysis}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: